---
title: "Analysis"
author: "Jessie Huang"
date: "2024-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fiftystater)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(betareg)
```

# Analysis

# Data Sources

```{r}
data_bmi <- read.csv('crc/bmi.csv', check.names = FALSE)
data_npo <- read.csv('crc/nutrition-physicalactivity-obesity.csv', check.names = FALSE)
data_physact <- read.csv('crc/physicalactivity.csv', check.names = FALSE)
```

```{r}
data('fifty_states')
colnames(fifty_states)[6] <- 'state_desc'
```

```{r}
head(data_bmi)
# EO-CRC rising in all races (especially White and Black) and males
risk_bmi_groups <- c('Obese (BMI 30.0 - 99.8)')
demographics <- c('18-24', '25-34', '35-44', '45-54', 'Black, non-Hispanic', 'White, non-Hispanic', 'American Indian or Alaskan Native, non-Hispanic', 'Asian, non-Hispanic', 'Native Hawaiian or other Pacific Islander, non-Hispanic', 'Multiracial, non-Hispanic', 'Other, non-Hispanic', 'Hispanic', 'Female', 'Male')
race <- c('Black, non-Hispanic', 'White, non-Hispanic', 'American Indian or Alaskan Native, non-Hispanic', 'Asian, non-Hispanic', 'Native Hawaiian or other Pacific Islander, non-Hispanic', 'Multiracial, non-Hispanic', 'Other, non-Hispanic', 'Hispanic')
age <- c('18-24', '25-34', '35-44', '45-54')
sex <- c('Female', 'Male')

age_levels <- c('18 - 24', '25 - 34', '35 - 44', '45 - 54')
sex_levels <- c('Female', 'Male')
race_levels <- c("2 or more races", "American Indian/Alaska Native", "Asian", "Hawaiian/Pacific Islander", "Hispanic", "Non-Hispanic Black", "Non-Hispanic White", "Other")
```

# Obesity

## Data Cleaning

```{r}
# data_bmi %>% filter(is.na(Data_value))
data_bmi_clean <- data_bmi %>% 
  select(Year, Locationabbr, Locationdesc, Break_Out, Response, Data_value) %>%
  filter(! is.na(Data_value))

data_bmi_viz <- data_bmi %>% 
  select(Year, Locationabbr, Locationdesc, Break_Out, Response, Data_value) %>%
  replace(is.na(.), 0)
head(data_bmi_clean)
```

```{r}
colnames(data_bmi_clean) <- c('Year', 'state_cd', 'state_desc', 'demographic', 'BMI', 'CrudePrevalence')
colnames(data_bmi_viz) <- c('Year', 'state_cd', 'state_desc', 'demographic', 'BMI', 'CrudePrevalence')

data_bmi_clean$CrudePrevalence <- ifelse(data_bmi_clean$CrudePrevalence!=0.0, data_bmi_clean$CrudePrevalence/100, 0)
data_bmi_clean$Year <- as.factor(data_bmi_clean$Year)
data_bmi_clean$state_cd <- as.factor(data_bmi_clean$state_cd)
data_bmi_clean$demographic <- as.factor(data_bmi_clean$demographic)
data_bmi_clean$BMI <- as.factor(data_bmi_clean$BMI)
data_bmi_clean$state_desc <- tolower(data_bmi_clean$state_desc)

data_bmi_viz$CrudePrevalence <- ifelse(data_bmi_viz$CrudePrevalence!=0.0, data_bmi_viz$CrudePrevalence/100, 0)
data_bmi_viz$Year <- as.factor(data_bmi_viz$Year)
data_bmi_viz$state_cd <- as.factor(data_bmi_viz$state_cd)
data_bmi_viz$demographic <- as.factor(data_bmi_viz$demographic)
data_bmi_viz$BMI <- as.factor(data_bmi_viz$BMI)
data_bmi_viz$state_desc <- tolower(data_bmi_viz$state_desc)
```


```{r}
data_bmi_risk <- data_bmi_clean %>% 
  filter(demographic %in% demographics & BMI %in% risk_bmi_groups)

data_bmi_age_risk <- data_bmi_risk %>% 
  filter(demographic %in% age & BMI %in% risk_bmi_groups)

data_bmi_sex_risk <- data_bmi_risk %>% 
  filter(demographic %in% sex & BMI %in% risk_bmi_groups)

data_bmi_race_risk <- data_bmi_risk %>% 
  filter(demographic %in% race & BMI %in% risk_bmi_groups & CrudePrevalence != 0)
```

```{r}
hist(data_bmi_age_risk$CrudePrevalence)
hist(data_bmi_sex_risk$CrudePrevalence)
hist(data_bmi_race_risk$CrudePrevalence)
```

## Age Model

```{r}
# finding demo's with a tendency toward obesity
# year by year all age groups report obese weight range by significant amount
# 45-54's in year 2021 was significant year of increase
mod.fit.age <- lm(CrudePrevalence ~ Year + demographic + Year*demographic, data=data_bmi_age_risk)
summary(mod.fit.age)
plot(mod.fit.age)
```


## Sex Model

```{r}
# finding demo's with a tendency toward obesity
# year by year less males report obese weight range by significant amount than females, especially 2021 and 2022
mod.fit.sex <- lm(CrudePrevalence ~ Year + demographic + Year*demographic, data=data_bmi_sex_risk)
summary(mod.fit.sex)
plot(mod.fit.sex)
```

## Race Model

### Year by Year
```{r}
# year by year 
# unclear, lot of missing data throughout the year of races, removing year for overall picture then. 
mod.fit.race <- lm(CrudePrevalence ~ Year + demographic + Year*demographic, data=data_bmi_race_risk)
summary(mod.fit.race)
plot(mod.fit.race)
```

# Only Race
```{r}
# remove year for clearer picture
# american indian/alaskan native, black, hispanic, multiracial, and white increase in obesity weight range report by significant amount
# others are decrease (but i suspect it's from lack of data versus actual reporting)
mod.fit.race2 <- lm(CrudePrevalence ~ demographic, data=data_bmi_race_risk)
summary(mod.fit.race2)
plot(mod.fit.race2)
```

## Visualizations

```{r}
# filter to most recent year for most recent behavior
data_bmi_clean_2022 <- data_bmi_viz %>% 
  filter(Year==2022)
head(data_bmi_clean_2022)
```

```{r, warning=FALSE}
risk_demo <- c('American Indian or Alaskan Native, non-Hispanic', 'Black, non-Hispanic', 'Hispanic', 'Multiracial, non-Hispanic', 'White, non-Hispanic')
data_risk_demo <- data_bmi_clean_2022 %>%
    group_by(state_desc, demographic) %>%
    filter(demographic %in% risk_demo & BMI %in% risk_bmi_groups) %>%
    arrange(desc(CrudePrevalence))
head(data_risk_demo)

# joins
data_risk_demo_geo <- left_join(fifty_states,data_risk_demo, by='state_desc')
head(data_risk_demo_geo)
```

```{r, fig.height=6}
ggplot(data_risk_demo_geo, aes(x=reorder(state_desc, CrudePrevalence), y=CrudePrevalence, color=demographic)) +
  geom_point() +
  geom_segment(aes(x=state_desc, xend=state_desc, y=0, yend=CrudePrevalence)) +
  coord_flip() +
  theme_classic() +
  # ggtitle('Obesity Prevalence by Race and US State (2022)') + 
  labs(x = 'State', y = 'Crude Prevalence %', color='Demographic') 
```

# Nutrition

## Data Cleaning
Not much data, will have to go off who isn't eating daily fruits and vegetables....
```{r}
data_nutrition <- data_npo %>% filter(Class == 'Fruits and Vegetables')

data_nutrition_filtered <- data_nutrition %>% 
  select(YearEnd, LocationAbbr, LocationDesc,
         `Age(years)`,Gender,`Race/Ethnicity`,
         Question, Data_Value) %>%
  filter(! is.na(Data_Value))

data_nutrition_viz <- data_nutrition %>% 
  select(YearEnd, LocationAbbr, LocationDesc,
         `Age(years)`,Gender,`Race/Ethnicity`,
         Question, Data_Value) %>%
  replace(is.na(.), 0)
head(data_nutrition_filtered)
```


```{r}
colnames(data_nutrition_filtered) <- c('Year', 'state_cd', 'state_desc', 'age', 'sex', 'race_ethnicity', 'diet_cat', 'percent')
colnames(data_nutrition_viz) <- c('Year', 'state_cd', 'state_desc', 'age', 'sex', 'race_ethnicity', 'diet_cat', 'percent')

data_nutrition_filtered$state_desc <- tolower(data_nutrition_filtered$state_desc)
data_nutrition_filtered$Year <- as.factor(data_nutrition_filtered$Year)
data_nutrition_filtered$state_cd <- as.factor(data_nutrition_filtered$state_cd)
data_nutrition_filtered$state_desc <- as.factor(data_nutrition_filtered$state_desc)
data_nutrition_filtered$age <- as.factor(data_nutrition_filtered$age)
data_nutrition_filtered$sex <- as.factor(data_nutrition_filtered$sex)
data_nutrition_filtered$race_ethnicity <- as.factor(data_nutrition_filtered$race_ethnicity)
data_nutrition_filtered$diet_cat <- as.factor(data_nutrition_filtered$diet_cat)
data_nutrition_filtered$percent <- ifelse(data_nutrition_filtered$percent!=0.0, data_nutrition_filtered$percent/100, 0.0)

data_nutrition_viz$state_desc <- tolower(data_nutrition_viz$state_desc)
data_nutrition_viz$Year <- as.factor(data_nutrition_viz$Year)
data_nutrition_viz$state_cd <- as.factor(data_nutrition_viz$state_cd)
data_nutrition_viz$state_desc <- as.factor(data_nutrition_viz$state_desc)
data_nutrition_viz$age <- as.factor(data_nutrition_viz$age)
data_nutrition_viz$sex <- as.factor(data_nutrition_viz$sex)
data_nutrition_viz$race_ethnicity <- as.factor(data_nutrition_viz$race_ethnicity)
data_nutrition_viz$diet_cat <- as.factor(data_nutrition_viz$diet_cat)
data_nutrition_viz$percent <- ifelse(data_nutrition_viz$percent!=0.0, data_nutrition_viz$percent/100, 0.0)

head(data_nutrition_viz)
```

## Vegetable 

```{r}
head(data_nutrition_filtered)

data_veg <- data_nutrition_filtered %>% 
  filter(diet_cat=='Percent of adults who report consuming vegetables less than one time daily')

data_fruit <- data_nutrition_filtered %>% 
  filter(diet_cat=='Percent of adults who report consuming fruit less than one time daily')

data_veg_viz <- data_nutrition_viz %>% 
  filter(diet_cat=='Percent of adults who report consuming vegetables less than one time daily')

data_fruit_viz <- data_nutrition_viz %>% 
  filter(diet_cat=='Percent of adults who report consuming fruit less than one time daily')
```

```{r}
# veg
data_veg_age <- data_veg %>% 
  filter(age %in% age_levels)
data_veg_sex <- data_veg %>% 
  filter(sex %in% sex_levels)
data_veg_race <- data_veg %>% 
  filter(race_ethnicity %in% race_levels)

# fruit
data_fruit_age <- data_fruit %>% 
  filter(age %in% age_levels)
data_fruit_sex <- data_fruit %>% 
  filter(sex %in% sex_levels)
data_fruit_race <- data_fruit %>% 
  filter(race_ethnicity %in% race_levels)
```

```{r}
hist(data_veg_age$percent)
hist(data_veg_sex$percent)
hist(data_veg_race$percent)
```
```{r}
hist(data_fruit_age$percent)
hist(data_fruit_sex$percent)
hist(data_fruit_race$percent)
```

## Age Model (Veg)

Going with beta distribution
- response variable distribution
- response is continuous probability between 0 and 1
- link: logit

interpret coefficients as odds:
https://stats.stackexchange.com/questions/442789/interpretation-of-the-beta-regression-coefficients-with-logit-link-used-to-analy

```{r, warning=FALSE}
# more younger not eating at least 1 fruit or 1 veg a day
# Fit beta regression model
mod.fit.veg.age <- betareg(percent ~ Year + age + Year*age, link = "logit", data = data_veg_age)
summary(mod.fit.veg.age)
```

```{r}
plot(mod.fit.veg.age, which=1) # check link appropriateness... constant variance here
plot(mod.fit.veg.age, which=2) # checking for outliers
# plot(mod.fit.veg.age, which=4) # minor heteroskedasticity
hist(mod.fit.veg.age$residuals) # wouldn't expect perfect normal distribution with beta distribution
qqnorm(residuals(mod.fit.veg.age, type = "sweighted2"))
qqline(residuals(mod.fit.veg.age, type = "sweighted2"))
```

## Sex Model (Veg)

```{r, warning=FALSE}
# more male not eating at least 1 fruit or 1 veg a day
head(data_veg_sex)
mod.fit.veg.sex <- betareg(percent ~ Year + sex + Year*sex, link = "logit", data = data_veg_sex)
summary(mod.fit.veg.sex)
```

```{r}
plot(mod.fit.veg.sex, which=1) # check link appropriateness... constant variance here
plot(mod.fit.veg.sex, which=2) # checking for outliers
# plot(mod.fit.nutrition.sex, which=4) # minor heteroskedasticity
hist(mod.fit.veg.sex$residuals) # wouldn't expect perfect normal distribution with beta distribution
qqnorm(residuals(mod.fit.veg.sex, type = "sweighted2"))
qqline(residuals(mod.fit.veg.sex, type = "sweighted2"))
```

## Race Model (Veg)

```{r, warning=FALSE}
# less likely to respond to not eat at least 1 fruit or 1 veg a day: asian, white, other
# more likely to respond to not eat at least 1 fruit or 1 veg a day: 2 or more races, american indian or alaska native, hawaiian/pacific islander, hispanic, black
mod.fit.veg.race <- betareg(percent ~ Year + race_ethnicity + Year*race_ethnicity, link = "logit", data = data_veg_race)
summary(mod.fit.veg.race)
```

```{r}
plot(mod.fit.veg.race, which=1) # check link appropriateness... constant variance here
plot(mod.fit.veg.race, which=2) # checking for outliers
# plot(mod.fit.veg.race, which=4) # minor heteroskedasticity
hist(mod.fit.veg.race$residuals) # wouldn't expect perfect normal distribution with beta distribution
qqnorm(residuals(mod.fit.veg.race, type = "sweighted2"))
qqline(residuals(mod.fit.veg.race, type = "sweighted2"))
```

## Age Model (Fruit)

```{r, warning=FALSE}
# more younger not eating at least 1 fruit or 1 veg a day
# Fit beta regression model
mod.fit.fruit.age <- betareg(percent ~ Year + age + Year*age, link = "logit", data = data_fruit_age)
summary(mod.fit.fruit.age)
```

```{r}
plot(mod.fit.fruit.age, which=1) # check link appropriateness... constant variance here
plot(mod.fit.fruit.age, which=2) # checking for outliers
# plot(mod.fit.veg.race, which=4) # minor heteroskedasticity
hist(mod.fit.fruit.age$residuals) # wouldn't expect perfect normal distribution with beta distribution
qqnorm(residuals(mod.fit.fruit.age, type = "sweighted2"))
qqline(residuals(mod.fit.fruit.age, type = "sweighted2"))
```


## Sex Model (Fruit)

```{r, warning=FALSE}
# more younger not eating at least 1 fruit or 1 veg a day
# Fit beta regression model
mod.fit.sex.age <- betareg(percent ~ Year + sex + Year*sex, link = "logit", data = data_fruit_sex)
summary(mod.fit.sex.age)
```

```{r}
plot(mod.fit.fruit.age, which=1) # check link appropriateness... constant variance here
plot(mod.fit.fruit.age, which=2) # checking for outliers
# plot(mod.fit.veg.race, which=4) # minor heteroskedasticity
hist(mod.fit.fruit.age$residuals) # wouldn't expect perfect normal distribution with beta distribution
qqnorm(residuals(mod.fit.fruit.age, type = "sweighted2"))
qqline(residuals(mod.fit.fruit.age, type = "sweighted2"))
```

## Race Model (Fruit)

```{r, warning=FALSE}
# more younger not eating at least 1 fruit or 1 veg a day
# Fit beta regression model
mod.fit.fruit.race <- betareg(percent ~ Year + race_ethnicity + Year*race_ethnicity, link = "logit", data = data_fruit_race)
summary(mod.fit.fruit.race)
```

```{r}
plot(mod.fit.fruit.race, which=1) # check link appropriateness... constant variance here
plot(mod.fit.fruit.race, which=2) # checking for outliers
# plot(mod.fit.veg.race, which=4) # minor heteroskedasticity
hist(mod.fit.fruit.race$residuals) # wouldn't expect perfect normal distribution with beta distribution
qqnorm(residuals(mod.fit.fruit.race, type = "sweighted2"))
qqline(residuals(mod.fit.fruit.race, type = "sweighted2"))
```


## Visualizations

```{r, warning=FALSE}
# 2019 most recent data for diet questions
data_nutrition_2019 <- data_nutrition_viz %>% 
  filter(Year == 2019 & state_cd != 'US')

# veg
data_veg_2019 <- data_nutrition_2019 %>% 
  filter(diet_cat == 'Percent of adults who report consuming vegetables less than one time daily')

data_veg_geo_2019 <- left_join(fifty_states,data_veg_2019, by='state_desc')

data_veg_geo_age <- data_veg_geo_2019 %>%
  filter(age %in% age_levels)
data_veg_geo_sex <- data_veg_geo_2019 %>%
  filter(sex %in% sex_levels)
data_veg_geo_race <- data_veg_geo_2019 %>%
  filter(race_ethnicity %in% race_levels)

# fruit
data_fruit_2019 <- data_nutrition_2019 %>% 
  filter(diet_cat == 'Percent of adults who report consuming fruit less than one time daily')
data_fruit_geo_2019 <- left_join(fifty_states,data_fruit_2019, by='state_desc')

data_fruit_geo_age <- data_fruit_geo_2019 %>%
  filter(age %in% age_levels)
data_fruit_geo_sex <- data_fruit_geo_2019 %>%
  filter(sex %in% sex_levels)
data_fruit_geo_race <- data_fruit_geo_2019 %>%
  filter(race_ethnicity %in% race_levels)
```

### Age (Veg)

```{r}
ggplot() + 
  geom_polygon(data=data_veg_geo_age, aes(x=long, y=lat, group = group, fill=percent), color='white') +
  scale_size(name="") + 
  facet_wrap(age~.)+
  theme_void() +
  labs(fill = 'Percent %')
```

### Sex (Veg)

```{r, fig.width=5, fig.height=5}
ggplot() + 
  geom_polygon(data=data_veg_geo_sex, aes(x=long, y=lat, group = group, fill=percent), color='white') +
  scale_size(name="") + 
  facet_grid(sex~.)+
  theme_void() +
  labs(fill = 'Percent %')
```

### Race (Veg)

```{r}
ggplot() + 
  geom_polygon(data=data_veg_geo_race, aes(x=long, y=lat, group = group, fill=percent), color='white') +
  scale_size(name="") + 
  facet_wrap(race_ethnicity~.)+
  theme_void() +
  labs(fill = 'Percent %')
```

### Age (Fruit)

```{r}
ggplot() + 
  geom_polygon(data=data_fruit_geo_age, aes(x=long, y=lat, group = group, fill=percent), color='white') +
  scale_size(name="") + 
  facet_wrap(age~.)+
  theme_void() +
  labs(fill = 'Percent %')
```

### Sex (Fruit)

```{r, fig.width=5, fig.height=5}
ggplot() + 
  geom_polygon(data=data_fruit_geo_sex, aes(x=long, y=lat, group = group, fill=percent), color='white') +
  scale_size(name="") + 
  facet_grid(sex~.)+
  theme_void() +
  labs(fill = 'Percent %')
```

### Race (Fruit)

```{r}
ggplot() + 
  geom_polygon(data=data_fruit_geo_race, aes(x=long, y=lat, group = group, fill=percent), color='white') +
  scale_size(name="") + 
  facet_wrap(race_ethnicity~.)+
  theme_void() +
  labs(fill = 'Percent %')
```

# Fitness

## Data Cleaning
CDC Guidelines:
Adults (18-64 years): At least 150 minutes a week of moderate intensity activity such as brisk walking. At least 2 days a week of activities that strengthen muscles. Aim for the recommended activity level but be as active as you are able.

```{r}
data_physact_q1 <- data_physact %>% 
  filter(Question == 'Participated in enough Aerobic and Muscle Strengthening exercises to meet guidelines (variable calculated from one or more BRFSS questions)') %>%
  select(Year, Locationabbr, Locationdesc, Break_Out, Response, Data_value) %>% 
  filter(! is.na(Data_value))

data_physact_q1_viz <- data_physact %>% 
  filter(Question == 'Participated in enough Aerobic and Muscle Strengthening exercises to meet guidelines (variable calculated from one or more BRFSS questions)') %>%
  select(Year, Locationabbr, Locationdesc, Break_Out, Response, Data_value) %>% 
  replace(is.na(.), 0)
head(data_physact_q1_viz)
```

```{r}
colnames(data_physact_q1) <- c('Year', 'state_cd', 'state_desc', 'demographic', 'Response', 'CrudePrevalence')
colnames(data_physact_q1_viz) <- c('Year', 'state_cd', 'state_desc', 'demographic', 'Response', 'CrudePrevalence')

data_physact_q1$Year <- as.factor(data_physact_q1$Year)
data_physact_q1$state_cd <- as.factor(data_physact_q1$state_cd)
data_physact_q1$demographic <- as.factor(data_physact_q1$demographic)
data_physact_q1$Response <- as.factor(data_physact_q1$Response)
data_physact_q1$state_desc <- tolower(data_physact_q1$state_desc)
data_physact_q1$CrudePrevalence <- ifelse(data_physact_q1$CrudePrevalence!=0.0, data_physact_q1$CrudePrevalence/100, 0.0)

data_physact_q1_viz$Year <- as.factor(data_physact_q1_viz$Year)
data_physact_q1_viz$state_cd <- as.factor(data_physact_q1_viz$state_cd)
data_physact_q1_viz$demographic <- as.factor(data_physact_q1_viz$demographic)
data_physact_q1_viz$Response <- as.factor(data_physact_q1_viz$Response)
data_physact_q1_viz$state_desc <- tolower(data_physact_q1_viz$state_desc)
data_physact_q1_viz$CrudePrevalence <- ifelse(data_physact_q1_viz$CrudePrevalence!=0.0, data_physact_q1_viz$CrudePrevalence/100, 0.0)
```

```{r}
# filter out the NAs
data_physact_inactive <- data_physact_q1 %>% 
  filter(Response=='No')

data_physact_inactive_age <- data_physact_inactive %>% 
  filter(demographic %in% age)
data_physact_inactive_sex <- data_physact_inactive %>% 
  filter(demographic %in% sex)
data_physact_inactive_race <- data_physact_inactive %>% 
  filter(demographic %in% race)

head(data_physact_inactive_age)
```

```{r}
hist(data_physact_inactive_age$CrudePrevalence)
hist(data_physact_inactive_sex$CrudePrevalence)
hist(data_physact_inactive_race$CrudePrevalence)
```

## Age Model

```{r}
# all age groups not meeting CDC guidelines for fitness but gets worse with older groups
mod.fit.inactive.age <- lm(CrudePrevalence ~ Year + demographic + Year*demographic, data=data_physact_inactive_age)
summary(mod.fit.inactive.age)
plot(mod.fit.inactive.age)
```

## Sex Model

```{r}
# males more likely to abide to CDC guidelines for fitness
mod.fit.inactive.sex <- lm(CrudePrevalence ~ Year + demographic + Year*demographic, data=data_physact_inactive_sex)
summary(mod.fit.inactive.sex)
plot(mod.fit.inactive.sex)
```

## Race Model

```{r}
# more likely to abide to CDC guidelines for fitness: asian, pacific islander (i suspect lack of data for significant decrease), and other
# less likely to abide to CDC guidelines for fitness: american indian/alaskan native, black, hispanic, multiracial, white
mod.fit.inactive.race <- lm(CrudePrevalence ~ Year + demographic + Year+demographic, data=data_physact_inactive_race)
summary(mod.fit.inactive.race)
plot(mod.fit.inactive.race)
```

## Visualizations

```{r}
data_physact2 <- data_npo %>% 
  filter(Class == 'Physical Activity' & Question == 'Percent of adults who engage in no leisure-time physical activity')

# 2019 most recent data for activity questions
data_physact2_2019 <- data_physact2 %>% 
  filter(YearEnd == 2019) %>%
  select(LocationAbbr, LocationDesc, 'Age(years)', Gender, 'Race/Ethnicity', Question, Data_Value) %>%
  replace(is.na(.), 0)

colnames(data_physact2_2019) <- c('state_cd', 'state_desc', 'age_group', 'sex', 'race_ethnicity', 'fitness_category', 'percent')
data_physact2_2019$state_desc <- tolower(data_physact2_2019$state_desc)
head(data_physact2_2019)
```


```{r, warning=FALSE}

data_inactive_geo_2019 <- left_join(fifty_states,data_physact2_2019, by='state_desc')

data_inactive_geo_age <- data_inactive_geo_2019 %>%
  filter(age_group %in% age_levels)

data_inactive_geo_sex <- data_inactive_geo_2019 %>%
  filter(sex %in% sex_levels)

data_inactive_geo_race <- data_inactive_geo_2019 %>%
  filter(race_ethnicity %in% race_levels)
```


### Age

```{r}
ggplot() + 
  geom_polygon(data=data_inactive_geo_age, aes(x=long, y=lat, group = group, fill=percent), color='white') +
  scale_size(name="") + 
  facet_wrap(age_group~.)+
  theme_void() +
  labs(fill = 'Percent %')
```

### Sex

```{r, fig.height=5, fig.width=5}
ggplot() + 
  geom_polygon(data=data_inactive_geo_sex, aes(x=long, y=lat, group = group, fill=percent), color='white') +
  scale_size(name="") + 
  facet_grid(sex~.)+
  theme_void() +
  # ggtitle('Inactive Adults - Race') + 
  labs(fill = 'Percent %')
```

### Race

from obesity analysis:
Races: American Indian/Alaska Native, Black, non-Hispanic, Hispanic, Multiracial, non-Hispanic, White, non-Hispanic
risk_states <- c('OK', 'AL', 'VA', 'SD', 'WI', 'ME', 'WY', 'MS', 'WV', 'IN')
```{r}
ggplot() + 
  geom_polygon(data=data_inactive_geo_race, aes(x=long, y=lat, group = group, fill=percent), color='white') +
  scale_size(name="") + 
  facet_wrap(race_ethnicity~.)+
  theme_void() +
  labs(fill = 'Percent %')
```
